---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-")
```
ajive: Angle based Joint and Individual Variation Explained
---------------

**Author:** [Iain Carmichael](https://idc9.github.io/)<br/>
**License:** [MIT](https://opensource.org/licenses/MIT)

Additional documentation, examples and code revisions are coming soon. For questions, issues or feature requests please reach out to Iain: iain@unc.edu.


# Overview

Angle based Joint and Individual Variation Explained (AJIVE) is a dimensionality reduction algorithm for the multi-block setting i.e. $K$ different data matrices, with the same set of observations and (possibly) different numbers of variables. **AJIVE finds *joint* modes of variation which are common to all $K$ data blocks as well as modes of *individual* variation which are specific to each block.** For a detailed discussion of AJIVE see [Angle-Based Joint and Individual Variation Explained](https://arxiv.org/pdf/1704.02060.pdf).

A python version of this package can be found [**here**](https://github.com/idc9/py_jive).

# Installation

The `ajive` package is currently available using devtools
```{r, eval=FALSE}
# install.packages('devtools')
devtools::install_github("idc9/r_jive")
```

# Example
Consider the following two block toy example: the first block has 200 observations (rows) and 100 variables; the second block has the same set of 200 observations and 500 variables (similar to Figure 2 of the AJIVE paper).
```{r, fig.width=8, fig.height=4}
library(ajive)
blocks <- sample_toy_data(n=200, dx=100, dy=500)

data_blocks_heatmap(blocks, show_color_bar=FALSE)
```

After selecting the initial signal ranks we can compute the AJIVE decomposition using the `ajive` function.
```{r}
initial_signal_ranks <- c(2, 2) # set by looking at scree plots
jive_decomposition <- ajive(blocks, initial_signal_ranks)
```


The heatmap below shows that AJIVE separates the joint and individual signals for this toy data set.
```{r, fig.width=8, fig.height=16}
decomposition_heatmaps(blocks, jive_decomposition)
```


Using notation from Section 3 of the [AJIVE paper](https://arxiv.org/pdf/1704.02060.pdf) (where *u* means scores and *v* means loadings) we can get the jive data out as follows
```{r}
# common normalized scores
jive_decomposition$joint_scores

# Full matrix representation of the joint signal for the first block
jive_decomposition[[1]][['joint']][['full']]

# joint block specific scores for the first block
jive_decomposition[[1]][['joint']][['u']]

# joint block specific loadings for the first block
jive_decomposition[[1]][['joint']][['v']]

# individual block specific scores for the second block
jive_decomposition[[2]][['individual']][['u']]
```
# Help and Support

Additional documentation, examples and code revisions are coming soon. For questions, issues or feature requests please reach out to Iain: iain@unc.edu. 

#### Documentation

The source code is located on github: [https://github.com/idc9/r_jive](https://github.com/idc9/r_jive). Currently the best math reference is the [AJIVE paper](https://arxiv.org/pdf/1704.02060.pdf).

#### Testing

Testing is done using the [testthat](https://github.com/hadley/testthat) package.

#### Contributing

We welcome contributions to make this a stronger package: data examples, bug fixes, spelling errors, new features, etc. <!-- TODO: add a more CONTRIBUTING file with more detail -->

#### Citation

A [Journal of Statistical Software](https://www.jstatsoft.org/index) paper is hopefully coming soon...
