---
title: "AJIVE basic usage"
author: "Iain Carmichael"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette covers the basic functionality of the `ajive` package.

```{r}
library(ajive)
```

# Sample toy data


```{r, fig.height=4, fig.width=8}
blocks <- sample_toy_data(n=200, dx=100, dy=500)
```

First decide on the initial signal ranks.
```{r, fig.height=4, fig.width=8}
scree_plot_blocks(blocks)
```

Using the above scree plots we set the initial signal rank equal to 2 for both blocks.
```{r}
initial_signal_ranks <- c(2, 2)
```

Now we are ready to compute the JIVE decomposition
```{r, cache=TRUE}
ajive_output <- ajive(blocks, initial_signal_ranks,
                      n_wedin_samples = 100, n_rand_dir_samples = 100)
```



# accessing AJIVE output

You can access a number of the possible quantities of interest with the following functions

- `get_block_full`
- `get_common_normalized_scores`
- `get_block_scores`
- `get_block_loadings`
- `get_joint_rank`
- `get_individual_rank`

The user can alwasy access the AJIVE output directly from `ajive_output` which is just a list.

## Full matrix representation

The user can access the full JIVE decomposition matrices with `get_block_full` i.e.
```{r}
k <- 1
J <- get_block_full(ajive_output, k, type='joint')
I <- get_block_full(ajive_output, k, type='individual')
E <- get_block_full(ajive_output, k, type='noise')

# check X = J + I + E
norm(blocks[[k]] - (J + I + E))
```

## Block Specific Scores

```{r}
bss_1 <- get_block_scores(ajive_output, k, type='joint', normalized=FALSE)

```


```{r}
normalized_bss_1 <- get_block_scores(ajive_output, k, type='joint', normalized=TRUE)

```


## Common Normalized Scores

```{r}
cns <- get_common_normalized_scores(ajive_output)
```


## Individual Normalized Scores

```{r}
ins_1 <- get_block_scores(ajive_output, k, type='individual', normalized=TRUE)
```

## Individual loadings
```{r}
indiv_loadings_1 <- get_block_loadings(ajive_output, k, type='individual')
```


